---
title: "Assignment 1"
author: "Kevin Gardner"
date: "1/30/2022"
output: word_document
---
# ------------------------------------------------
# Following is the link to my GitHub account:
# https://github.com/Kgardner22/64060_-kgardner
# ------------------------------------------------


# 1. Download a dataset from the web. 
# My data source is Kaggle. Following are the details:
#     House Prices – Advanced Regression Techniques
#     Predict sales prices and practice feature engineering, RFs, and gradient boosting
# 
# https://www.kaggle.com/c/house-prices-advanced-regression-techniques/data


# 2. Import the dataset into R

```{r}
  House_Prices_train <- read.csv("C:/R/MyData/House_Prices_train.csv", header=TRUE)
```

# 3. Print out descriptive statistics for a selection of quantitative and categorical variables.

```{r}  
# The summary command will show a variety of descriptive statistics for each variable in the data set including the minimum, 1st quartile, median, mean, 3rd quartile, maximum values and if any NAs are present

  summary(House_Prices_train)
```

# Following are individual descriptive statistics for quantitative variables:

```{r}  
  mean(House_Prices_train$SalePrice)  # Mean Sale Price
```
```{r}  
  median(House_Prices_train$SalePrice) # Median Sale Price
```
```{r}  
  sd(House_Prices_train$SalePrice) # Standard Deviation of Sale Price
```
```{r}  
  min(House_Prices_train$SalePrice) # Minimum Sale Price
```
```{r}  
  max(House_Prices_train$SalePrice) # Maximum Sale Price
```

# Following are descriptive statistics for categorical variables:
```{r}  
  table(House_Prices_train$Street) # shows the frequency of homes located on gravel streets vs paved streets
```
```{r}  
  table(House_Prices_train$CentralAir) # shows the frequency of homes with and without central air
```
```{r}  
  table(House_Prices_train$CentralAir, House_Prices_train$Electrical) # cross classification of homes with and without central air (Y/N) and the type of electrical for the home (fuse box, electrical)
```

# To show the percentage of the frequency for each value in a specific categorical variable (such as SaleCondition)
```{r}
  table1 <- table(House_Prices_train$SaleCondition)
  prop.table(table1)
```  

# 4. Transform at least one variable. It doesn’t matter what the transformation is.
```{r}
  House_Prices_train$SalePrice_Transformed <- (House_Prices_train$SalePrice - mean(House_Prices_train$SalePrice))/sd(House_Prices_train$SalePrice)

  
# Create a new variable for total square feet (TotalLivingSF) which is the square footage of the 1st and 2nd floor combined:
  
  House_Prices_train$TotalLivingSF <- (House_Prices_train$X1stFlrSF + House_Prices_train$X2ndFlrSF)
```

# 5. Plot at least one quantitative variable, and one scatterplot.
```{r}
# Show histogram of SalePrice
  
  hist(House_Prices_train$SalePrice)
```
```{r}
# We could also add color to the Histogram to improve the visualization

  hist(House_Prices_train$SalePrice, col = blues9)
```

```{r}
# Show scatterplot of sales price (SalePrice) to total square footage (TotalLivingSF):
  
  plot(House_Prices_train$SalePrice, House_Prices_train$TotalLivingSF)
```  
```{r}
# We can add color to the scatterplot as well to improve the visualization

  plot(House_Prices_train$SalePrice, House_Prices_train$TotalLivingSF, col = blues9)
```


