---
title: "Assignment 2"
author: "Kevin Gardner"
date: "2/15/2022"
output:
  word_document: default
  html_document: default
---

# ------------------------------------------------
# Following is the link to my GitHub account:
# https://github.com/Kgardner22/64060_-kgardner
# ------------------------------------------------

Import and Prepare Data:
  
Import the UniversalBank.csv file  

```{r}

UniversalBank <- read.table('C:/R/MyData/UniversalBank.csv', header = T, sep = ',') 

summary(UniversalBank)

```
Create a copy of the original data file to preserve

```{r}

Original_File <- UniversalBank

```

REQUIREMENT 1:

Transform categorical predictors with more than two categories into dummy variables FIRST.
Need to do this for 'Education' and 'Personal.Loan'.

```{r}

library(caret)
library(class)

# Remove unnecessary attributes
UniversalBank$ID <- NULL
UniversalBank$ZIP.Code <- NULL

# Transform to factors
UniversalBank$Education=as.factor(UniversalBank$Education)
UniversalBank$Personal.Loan=as.factor(UniversalBank$Personal.Loan)

# Use dummyVars function to create a model
dummies <- dummyVars(Personal.Loan ~ ., data = UniversalBank)
UniversalBank_dummy <- as.data.frame(predict(dummies, newdata = UniversalBank))

```

Normalize the data (I'm using z-score scaling as input method)

```{r}

Norm_model <- preProcess(UniversalBank_dummy, method = c("center", "scale"))
UniversalBank_norm <- predict(Norm_model, UniversalBank_dummy)
summary(UniversalBank_norm)

```
Add back in the target variable (Personal.Loan)

```{r}

UniversalBank_norm$Personal.Loan <- UniversalBank$Personal.Loan

```

We need to divide the data into training (60%) and validation (40%) sets

```{r}

Train_Index <- createDataPartition(UniversalBank$Personal.Loan, p=0.6, list = FALSE)
Train.df <- UniversalBank_norm[Train_Index,]
Validation.df <- UniversalBank_norm[-Train_Index,]

```

Create data frame with values to predict
Age = 40, Experience = 10, Income = 84, Family = 2, CCAvg = 2, Education_1 = 0, Education_2 = 1, Education_3 = 0, Mortgage = 0, Securities Account = 0, CD Account = 0, Online = 1, and Credit Card = 1.

```{r}

To_Predict <- data.frame (Age=40, Experience=10, Income=84, Family=2, CCAvg=2, Education.1=0, Education.2=1, Education.3=0, Mortgage=0, Securities.Account=0, CD.Account=0, Online=1, CreditCard=1)

print(To_Predict)

```

Normalize this new record (To_Predict) using the same model we applied to the original dataset

```{r}

To_Predict_norm <- predict(Norm_model, To_Predict)
print(To_Predict_norm)

```

Use k-NN function to make the prediction.

Perform a k-NN classification with all predictors EXCEPT ID and Zip_Code using k=1.
Specify success class as 1 (loan acceptance) and use the default cutoff value of 0.5.

```{r}

Prediction <- knn(train = Train.df[,1:13],
                  test = To_Predict_norm[,1:13],
                  cl=Train.df$Personal.Loan,
                  k=1)
print(Prediction)

```
ANSWER - REQUIREMENT 1:

As shown in the above results (with k=1), the prediction for this observation is that Personal.Loan = 0 meaning, this individual is predicted to NOT accept the personal loan being offered.


REQUIREMENT 2:

What is a choice of k that balances between overfitting and ignoring the predictor information?

```{r}

set.seed(123)

fitControl <- trainControl(method = "repeatedcv", number = 3, repeats = 2)

searchGrid <- expand.grid(k = 1:15)

Knn.model <- train(Personal.Loan ~ .,
                   data = Train.df,
                   method = 'knn',
                   tuneGrid = searchGrid,
                   trControl = fitControl,)

Knn.model

```
ANSWER - REQUIREMENT 2:

The above Knn.model indicates that the k value with the highest accuracy is a value of k=3


REQUIREMENT 3:

Show the confusion matrix for the validation data that results from using the best k.

Use the predict function of the caret package to make predictions on the validation set.

```{r}

predictions <- predict(Knn.model, Validation.df)

```

Compare predictions from the Knn.model to the actual Personal.Loan labels in the validation set to compute the confusion matrix

```{r}

confusionMatrix(predictions, Validation.df$Personal.Loan)

```

REQUIREMENT 4:

Consider the following customer: Age = 40, Experience = 10, Income = 84, Family = 2, CCAvg = 2, Education_1 = 0, Education_2 = 1, Education_3 = 0, Mortgage = 0, Securities Account = 0, CD Account = 0, Online = 1 and Credit Card = 1. Classify the customer using the best k.

```{r}

# The independent variables specified in this requirement are the same as those specified in requirement 1. Therefore, this data.frame is already build and normalized (To_Predict_norm).

print(To_Predict)

```

Using the normalized prediction file (To_Predict_norm), we will use the Knn.model to predict using the best k value (k=3)

```{r}

predict(Knn.model, To_Predict_norm)

```
ANSWER - REQUIREMENT 4:

Using the best k value, the above results (with k=3) is predicting this observation will have Personal.Loan = 0 meaning, this individual is predicted to NOT accept the personal loan being offered.


